upwell_IDX

### 2020-04-25 SST_patterns.Rmd
update

# # All the code below can be viewed in upwell_IDX.Rmd
# # Next to run upwelling index
# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
# load("Data/UI_angle.RData")
# 
upwelling <- UI_angle %>%
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t),
  #        max_t = max(t)) %>%
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>%
  select(-data, -clim) %>%
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance)

@@

@@

@@

save(combined_freq_spread, )

>>>>

# # All the code below can be viewed in upwell_IDX.Rmd
# # Next to run upwelling index
# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)

# load("Data/UI_angle.RData")
load("Data_coast_angle/UI_angle.RData")

upwelling <- UI_angle %>%
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t),
  #        max_t = max(t)) %>%
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-07-09", "2017-12-21")), # creating a column of climatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>%
  select(-data, -clim) %>%
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance)

# Now applying the Upwelling func
# NB: This only pulls out the event results and not the climatology results
# This is done to keep the output tidy because group_modify() may only create data.frame type outputs
# To create a list output one would use group_map(),
# but this then loses the labels of which sites etc. the results belong to
detect_event_custom <- function(df){
  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 1, coldSpells = T)$event # 1 or 3?
  return(res)
}
# 
ts2clm_custom <- function(df){
  # The climatology base period used here is up for debate...
  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("2010-06-23", "2014-06-30"))
  return(res)
}
# 
# Calculate the upwelling event metrics
upwelling_detect_event <- function(df){
  upwell_base <- df %>%
    dplyr::rename(t = date) %>%
    group_by(site, product, heading, distance, lon, lat) %>%
    group_modify(~ts2clm_custom(.x)) %>%
    left_join(upwelling, by = c("site", "t")) %>%
    filter(!is.na(exceedance)) %>%
    group_by(site, product, heading, distance, lon, lat) %>%
    group_modify(~detect_event_custom(.x))
  }
# 
OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
#save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
save(OISST_upwell_base, file = "Data_coast_angle/OISST_upwell_base.RData")
CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
save(CMC_upwell_base, file = "Data_coast_angle/CMC_upwell_base.RData")
MUR_upwell_base <- upwelling_detect_event(df = Final)
save(MUR_upwell_base, file = "Data_coast_angle/MUR_upwell_base.RData")

# Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
SACTN_US <- SACTN_US %>%
  filter(site %in% selected_sites)

@@

save(SACTN_upwell_base, file = "Data_coast_angle/SACTN_upwell_base.RData")
```

@@

```{r}
# load("Data/OISST_upwell_base.RData")
# load("Data/CMC_upwell_base.RData")
# load("Data/SACTN_upwell_base.RData")
# load("Data/MUR_upwell_base.RData")
# load("Data/G1SST_upwell_base.RData")

load("Data_coast_base/OISST_upwell_base.RData")
load("Data_coast_base/CMC_upwell_base.RData")
load("Data_coast_base/SACTN_upwell_base.RData")
load("Data_coast_base/MUR_upwell_base.RData")
load("Data_coast_base/G1SST_upwell_base.RData")

@@

### 2020-04-25 upwell_IDX.Rmd
update

# RWS: Please figure out why some wind direction values are returned as NA
# RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
# RWS: Please look through these resources and make corrections accordingly:
# https://cran.r-project.org/web/packages/rWind/rWind.pdf
# https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r

wind_daily <- wind_data %>% 
  filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
  dplyr::group_by(temp_sites, date) %>%
  dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
                                                           template = 'geographics')),2),
                                    spw_circ = mean.circular(circular(speed, units = "degrees"))) # Direction using the circular function

save(wind_daily, file = "Data/wind_daily.RData") 
```

@@

```{r}
load("Data/wind_daily.RData") 
wind_daily <- wind_daily %>%
  dplyr::rename(sites = temp_sites) %>% 
  drop_na()
load("Data/site_list_sub.Rdata")
load("Data/SACTN_US.RData")
```

@@

```{r}
UI_angle <- wind_daily %>%  # Making reference to the wind daily data created above
  dplyr::rename(site = sites) %>%  # Renaming sites to site
  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to South Africa wide data column 2 and 21. by site 
  drop_na() %>% # removing the na values
  dplyr::rename(coast_angle = heading) %>% # Renaming the column heading to coast_angle
  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # applying the ui formula
  dplyr::rename(t = date) %>% 
  drop_na()#renaming date to t

@@

```{r}
# Loading the insitu temperature data along the wc
load("Data/SACTN_US.RData") # temperature data for all the sites with a 30year time series
# exceedance <-
#   function(data,
#            x = t,
#            y = temp,
#            threshold,
#            below = FALSE,
#            minDuration = 5,
#            joinAcrossGaps = TRUE,
#            maxGap = 2,
#            maxPadLength = 3)
# Upwelling temperature threshold 12.4 / 30th percentile
SACTN_upwell <- UI_angle %>% 
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t), 
  #        max_t = max(t)) %>% 
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-01-01", "2016-12-31")), # creating a column of climaatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>% 
  select(-data, -clim) %>% 
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance) # selecting only these variables
# Calculate quantiles of upwelling index

@@

# Detect event: 
detect_event_custom <- function(df){
  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)
  return(res)
}
# Calculate the upwelling event metrics
SACTN_upwell_base <- SACTN_US %>% 
  dplyr::rename(t = date) %>% 
  group_by(site) %>% 
  nest() %>% 
  # The climatology base period used here is up for debate...
  # The choice of the 30th percentile threshold also needs to be justified and sensitivty tested
  mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1995-01-01", "2004-12-31"))) %>%
  select(-data) %>% 
  unnest() %>%
  left_join(SACTN_upwell, by = c("site", "t")) %>%
  filter(!is.na(exceedance)) %>%
  group_by(site) %>% 
  # mutate(thresh = mean(seas, na.rm = T)-sd(seas, na.rm = T)) %>% # Manually set threshold to the mean
  # mutate(thresh = 10) %>% # Manually set threshold to a static value. Doesn't work across all sites
  # mutate(thresh = quantile(temp, 0.3, na.rm = T)) %>% # Manually set threshold to a single quantile
  nest() %>% 
  mutate(exceed = purrr::map(data, detect_event_custom)) %>% 
  select(-data) #%>% 
  # unnest() %>% 
  # filter(row_number() %% 2 == 0) %>% # Select event summary metrics
  # filter(row_number() %% 2 == 1) %>% # Select daily values
  # unnest()
# Unpack the event metric reults
SACTN_upwell_events <- SACTN_upwell_base %>% 
  unnest() %>%
  filter(row_number() %% 2 == 0) %>%
  unnest()
# save(SACTN_upwell_events, file = "Data/SACTN_upwell_events.RData")
# Unpack the daily climatology results
SACTN_upwell_clims <- SACTN_upwell_base %>% 
  unnest() %>%
  filter(row_number() %% 2 == 1) %>% 
  unnest()
# save(SACTN_upwell_clims, file = "Data/SACTN_upwell_clims.RData")
# The above chunk of code appears to work as expected.
# What needs to be done now is that the thresholds for duration and UI strength need to be justified/decided uopn. 
    # Min duration for an upwelling event - 1day and threshold -1
# What also needs to be decided is if we are interested in any upwelling results throughout the year,
# or only during the upwelling season, whenever that may be for each site.
# I think we have demonstrated the technical capability needed to answer the question of whether or not upwelling
# is changing over time.
# Now we really need to focus on which of these parameters need to best be tweaked to answer that question.
```

>>>>

# RWS: Please figure out why some wind direction values are returned as NA
# RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
# RWS: Please look through these resources and make corrections accordingly:
# https://cran.r-project.org/web/packages/rWind/rWind.pdfsave(, file = "Data_coastal_angle/wind_daily.RData")
# https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
# The data provided by SAWS already reads wind from the direction which it is blwing

wind_daily <- wind_data %>% 
  filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
  dplyr::group_by(temp_sites, date) %>%
  dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
                                                           template = 'geographics')),2),
                                    spw_circ = mean.circular(circular(speed, units = "degrees"))) # Direction using the circular function
# save(wind_daily, file = "Data_coast_angle/wind_daily.RData")

# library(openair)
# wind_daily_speed <- timeAverage(wind_data, avg.time = "day")
# # save(wind_daily, file = "Data/wind_daily.RData") 
# #Convert Wind Speed and Wind Direction to u/v Wind Components 
# # Then get the daily average and then convert it back?
# # https://rdrr.io/github/retostauffer/Rfoehnix/man/ddff2uv.html
# # Use with zoo
# 
# library("zoo")
# set.seed(100)
# Sys.setenv("TZ" = "UTC")
# ## Calculate dd/ff
# wind_data <- wind_data %>% 
#   rename(dd = dir,
#          ff = speed)
# 
# uv <- ddff2uv(wind_data)
# #blowing to
```

@@

```{r}
# load("Data/wind_daily.RData") 
load("Data_coast_angle/wind_daily.RData") 
wind_daily <- wind_daily %>%
  dplyr::rename(sites = temp_sites,
                mean_speed = spw_circ) %>% 
  drop_na()
load("Data/site_list_sub.Rdata")
load("Data/SACTN_US.RData")
```

@@

```{r}
UI_angle <- wind_daily %>%  # Making reference to the wind daily data created above
  dplyr::rename(site = sites) %>%  # Renaming sites to site
  #left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to South Africa wide data column 2 and 21. by site 
  drop_na() %>% # removing the na values
  #dplyr::rename(coast_angle = heading) %>% # Renaming the column heading to coast_angle
  mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # This is the rientation of the coastal angle for the west coast
  dplyr::rename(t = date) %>% 
  drop_na()#renaming date to t

# save(UI_angle, file = "Data_coast_angle/UI_angle.RData")

@@

```{r}
# Loading the insitu temperature data along the wc
load("Data/SACTN_US.RData") # temperature data for all the sites with a 30year time series
# exceedance <-
#   function(data,
#            x = t,
#            y = temp,
#            threshold,
#            below = FALSE,
#            minDuration = 5,
#            joinAcrossGaps = TRUE,
#            maxGap = 2,
#            maxPadLength = 3)
# Upwelling temperature threshold 12.4 / 30th percentile
SACTN_upwell <- UI_angle %>% 
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t), 
  #        max_t = max(t)) %>% 
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-07-09", "2017-12-21")), # creating a column of climaatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>% 
  select(-data, -clim) %>% 
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance) # selecting only these variables
# Calculate quantiles of upwelling index

@@

# Detect event: 
detect_event_custom <- function(df){
  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)
  return(res)
}
# Calculate the upwelling event metrics
SACTN_upwell_base <- SACTN_US %>% 
  dplyr::rename(t = date) %>% 
  group_by(site) %>% 
  nest() %>% 
  # The climatology base period used here is up for debate...
  # The choice of the 30th percentile threshold also needs to be justified and sensitivty tested
  mutate(clim = purrr::map(data, ts2clm, pctile = 25, climatologyPeriod = c("1992-07-09", "2005-10-19"))) %>%
  select(-data) %>% 
  unnest() %>%
  left_join(SACTN_upwell, by = c("site", "t")) %>%
  filter(!is.na(exceedance)) %>%
  group_by(site) %>% 
  # mutate(thresh = mean(seas, na.rm = T)-sd(seas, na.rm = T)) %>% # Manually set threshold to the mean
  # mutate(thresh = 10) %>% # Manually set threshold to a static value. Doesn't work across all sites
  # mutate(thresh = quantile(temp, 0.3, na.rm = T)) %>% # Manually set threshold to a single quantile
  nest() %>% 
  mutate(exceed = purrr::map(data, detect_event_custom)) %>% 
  select(-data) #%>% 
  # unnest() %>% 
  # filter(row_number() %% 2 == 0) %>% # Select event summary metrics
  # filter(row_number() %% 2 == 1) %>% # Select daily values
  # unnest()
# Unpack the event metric reults
SACTN_upwell_events <- SACTN_upwell_base %>% 
  unnest() %>%
  filter(row_number() %% 2 == 0) %>%
  unnest()
# save(SACTN_upwell_events, file = "Data/SACTN_upwell_events.RData")
# save(SACTN_upwell_events, file = "Data_coast_angle/SACTN_upwell_events.RData")
# Unpack the daily climatology results
SACTN_upwell_clims <- SACTN_upwell_base %>% 
  unnest() %>%
  filter(row_number() %% 2 == 1) %>% 
  unnest()
# save(SACTN_upwell_clims, file = "Data/SACTN_upwell_clims.RData")
# The above chunk of code appears to work as expected.
# What needs to be done now is that the thresholds for duration and UI strength need to be justified/decided uopn. 
    # Min duration for an upwelling event - 1day and threshold -1
# What also needs to be decided is if we are interested in any upwelling results throughout the year,
# or only during the upwelling season, whenever that may be for each site.
# I think we have demonstrated the technical capability needed to answer the question of whether or not upwelling
# is changing over time.
# Now we really need to focus on which of these parameters need to best be tweaked to answer that question.
```
